PMV: Which formulation should you use the ISO 7730:2005 or ASHRAE 55:2020?

### 2023-08-29 Manuscript/src/myacronyms.tex
paper: reviewing Stefano's comments

>>>>

    \acro{7730}[ISO~7730:2005]{ISO~7730:2005 is a thermal comfort standard developed by ISO}
    \acro{55}[ASHRAE~55:2023]{ASHRAE~55:2023 is a thermal comfort standard developed by ANSI and ASHRAE}
    \acro{pmv-ce}[PMV\textsubscript{CE}]{PMV calculated in accordance with the ASHRAE~55:2023}
    \acro{epmv}[ePMV]{Adjusted Predicted Mean Votes with Expectancy Factor}
    \acro{apmv}[aPMV]{Adaptive Predicted Mean Vote}
    \acro{pmvs}[PMV$_{SET}$]{PMV calculated using the SET temperature in place of $t_{db}$}
    \acro{pmvg}[PMV$_{Gagge}$]{PMV calculated using the two-node heat balance model}
    \acro{db2}[DB$_{comfort}$]{ASHRAE Global Thermal Comfort Database II}
    \acro{lowess}[LOWESS]{LOcally WEighted Scatterplot Smoothing}

### 2023-08-29 Manuscript/src/sections/conclusions.tex
paper: reviewing Stefano's comments

\section{Conclusions}\label{sec:conclusions}
The present study aimed to compare the accuracy of the \ac{pmv} model, as used in the \gls{7730} Standard, and its modified version, the \gls{pmv-ce} model, as employed in the \gls{55} Standard.
Our findings revealed notable disparities in the predictive accuracy of the \ac{pmv} and \gls{pmv-ce} models when compared against the \ac{tsv} available in the ASHRAE Global Thermal Comfort Database II v2.1 (Comfort DB).
These disparities shed light on the limitations and strengths of the \ac{pmv} model, enabling a more informed decision-making process for the selection of an appropriate formulation within National and International Standards.
The outcomes of this analysis have significant implications for researchers, practitioners, and policymakers involved in the field of thermal comfort assessment and standardization.

>>>>

\section{Conclusions}\label{sec:conclusions}
The present study aimed to compare the accuracy of the \ac{pmv} model, as used in the \gls{7730} Standard, and its modified version, the \ac{pmv-ce} model, as employed in the \gls{55} Standard.
Our findings revealed notable disparities in the predictive accuracy of the \ac{pmv} and \ac{pmv-ce} models when compared against the \ac{tsv} available in the ASHRAE Global Thermal Comfort Database II v2.1 (Comfort DB).
These disparities shed light on the limitations and strengths of the \ac{pmv} and \ac{pmv-ce} models, enabling a more informed decision-making process for the selection of an appropriate formulation within National and International Standards.
The outcomes of this analysis have significant implications for researchers, practitioners, and policymakers involved in the field of thermal comfort assessment and standardization.

### 2023-08-29 Manuscript/src/sections/introduction.tex
paper: reviewing Stefano's comments

\section{Introduction}\label{sec:introduction}

In 1970, Fanger~\cite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} Standard~\cite{iso7730} in its original form.
% todo also mention the other standards that use the PMV The PMV index has been widely used in national and international standards and guidance ASHRAE 55 [4], EN 15251 (new version EN 16798 [5]) [6], ISO 7730 [1], CIBSE Guide A [7] and GB/T 50785 [110] as  shown in  Table 5. This sentence is from Yao 2022.
The \ac{pmv} model has been widely adopted worldwide both among researchers and practitioners.
It is the most extensively used thermal comfort index for estimating the thermal sensation of a group of people sharing the same thermal environment.
A search within article title, abstract, and keyword on Scopus with the search term ``predicted mean vote'' returns approximately 1600 documents.
Out of those 1082 are scientific peer-review articles published in the research area of engineering, environmental science, social sciences, and energy.
This highlights the extensive adoption and use of this model among the scientific community.

To this date, the \ac{pmv} still remains to be the most utilized thermal comfort model despite the fact that several research papers have highlighted that the \ac{pmv} has a relatively low accuracy in predicting self-reported thermal sensation of people~\cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal,Humphreys2002,Doherty1988}.

To overcome this problem, the \gls{55} Standard uses a modified version of the original model developed by Fanger, the \gls{pmv-ce}~\cite{ashrae552020}.
The process used to calculate the \gls{pmv-ce} in the \gls{55} is summarized in the flowchart depicted in Figure~\ref{fig:flowchart_pmv_ce}.
In summary, when the \ac{v} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
The value of \ac{ce} is calculated using the \ac{set} equation and it is then subtracted from both the measured \ac{tdb} and \ac{tr}.
The resulting values become the new inputs into the \ac{pmv} model.
Since the \ac{ce} accounts for the convective heat losses from the person to its environment, thus, the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
While the other three input parameters (\ac{clo}, \ac{met}, and \ac{rh}) remain unchanged.
Consequently, for a given thermal environment, results from the two \ac{pmv} formulation differ when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
The rationale for this approach is that the original formulation of the \ac{pmv} model cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v} is higher than \qty{0.1}{\m\per\s}.
This method, however, has its own limitation, as follows.
Firstly, it should be noted that throughout the years the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552020}.
Secondly, calculating the \gls{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the two-node model.
Thirdly, we could not find any peer-reviewed scientific publication that described the benefits of using this model over the original \ac{pmv}.

@@

        \draw [arrow] (start) -- (dec1);
        \draw [arrow] (dec1) -- node[above,pos=0.3] {Yes} (pro1a);
        \draw [arrow] (dec1) -- node[right,pos=0.3] {No} (dec2);
        \draw [arrow] (pro1a) -- ($(pro1a.south)+(0,-0.5)$) -| (dec2);
        \draw [arrow] (dec2) -- node[above,pos=0.3] {Yes} (pro2a);
        \draw [arrow] (dec2) -- node[right,pos=0.3] {No} (dec3);
        \draw [arrow] (pro2a) -- ($(pro2a.south)+(0,-0.5)$) -| (dec3);
        \draw [arrow] (dec3) -- ($(dec3.west)$) -| node[above,pos=0.3] {No} (end1);
        \draw [arrow] (dec3) -- ($(dec3.east)$) -| node[above,pos=0.3] {Yes} (pro3);
        \draw [arrow] (pro3) -- (end2);
    \end{tikzpicture}
    \caption{Flowchart depicting the steps needed to calculate the PMV model in accordance with the ASHRAE 55--2020 Standard.}
    \label{fig:flowchart_pmv_ce}
\end{figure}

Choosing between the \ac{pmv} and \gls{pmv-ce} consequently is now a source of confusion for many practitioners, educators, and researchers worldwide since both models are widely adopted and incorporated in many national and international building codes worldwide.

\subsection{Other PMV formulations}\label{subsec:other-pmv-formulations}
To further complicate this issue, several other formulations of the \ac{pmv} model have been proposed worldwide.
Among the most notable formulations there are the: \gls{pmvs}~\cite{GaggeSET}, \gls{pmvg}~\cite{GaggeSET}, \gls{epmv}~\cite{Toftum2002}, and \ac{athb}~\cite{Schweiker2022}.
\mycite{Yao2022} provides a comprehensive review of different thermal comfort models and compares and describe some of the above-mentioned models.
Hence, we invite our readers to read \mycite{Yao2022} paper to learn more about some of \ac{pmv} the different variations that have been proposed.

%In 2002, Fanger, P. O. and Toftum, J. developed the \gls{epmv} an extension of the \ac{pmv}.
%The \gls{epmv} includes an expectancy factor introduced for use in non-air-conditioned buildings in warm climates.
%The authors argued that people adapt to the warm environment by decreasing their \ac{met} rate.
%They, consequently, propose to reduce the metabolic rate used to calculate the \ac{pmv} by \qty{6.7}{\percent} for every scale unit of \ac{pmv} above neutral.
%The resulting \ac{pmv} value is multiplied by the \ac{ef}.
%The \ac{ef} for non-air-conditioned buildings can be determined by the length of warm weather throughout the year and whether such buildings may be compared to many others in the region that are air-conditioned.
%If the temperature is hot all year or most of the year and there are no or few other air-conditioned buildings, \ac{ef} may be \num{0.5}, but if there are many other air-conditioned buildings, \ac{ef} may be \num{0.7}.
%For non-air-conditioned buildings in locations where the temperature is warm solely during the summer and no or few buildings have air-conditioning \ac{ef} is \numrange{0.7}{0.8}, whereas it is \numrange{0.8}{.9} in areas where there are numerous air-conditioned buildings.
%In regions with only brief periods of warm weather during the summer, the expectancy factor may be \numrange{0.9}{1}.
%This arbitrary and non-scientific definition of the expectancy factor make the \gls{epmv} difficult to calculate and non-generalizable.
%Consequently, the \gls{epmv} has not gained any widespread adoption.
%
%Yao, R. et al. (2009) developed the Adaptive Predicted Mean Vote Model~\gls{apmv} to account for occupants adaptation.
%They used a ``Black Box'' model to estimate an adaptive coefficient which is in turn use to calculate the \gls{apmv} using the following equation: $\mathrm{aPMV} = \mathrm{PMV}/(1+\lambda \mathrm{PMV})$
%The main issue with this approach is that the adaptive coefficient cannot be known a priori, but it is estimated based on the \ac{tsv} collected by a cohort of participants.
%This approach cannot be, therefore, generalizable and similarly to the \gls{epmv} model the \gls{apmv} is rarely used in thermal comfort research.
%

\subsection{Comparison of ISO and ASHRAE PMV formulations}\label{subsec:comparision-of-pmv-formulations}
To our knowledge, no previous study compared in detail the accuracy of the \ac{pmv} and \gls{pmv-ce} models which are currently used in the \gls{7730} and \gls{55} Standards.
In particular, we could not find any study which answered the questions: ``which \ac{pmv} formulation should you use the \gls{7730} or \gls{55}?''.

Some notable works include, the work from \mycite{Doherty1988} who evaluated the ability of the \ac{pmv} and \ac{set} models to predict several physiological variables (i.e., skin temperature, core temperature, and skin wettedness) under a wide range of still air environments and metabolic rates.
They concluded that the \ac{pmv} model is accurate for simulations of resting subjects but its accuracy decreases as a function of metabolic rate.
Humphreys and Nicol estimated the effects of measurement and formulation error on predicting thermal sensation using the \ac{pmv}~\cite{Humphreys2000}.
They used the ASHRAE Thermal Comfort database v1.0 and determined that the measurement error and the error introduced by the \ac{pmv} model formulation had a similar and non-negligible contribution.
Humphreys and Nicol also determined the validity of the \ac{pmv} for predicting comfort votes collected in field studies~\cite{Humphreys2002}.
They concluded that the \ac{pmv} range of applicability should be significantly reduced and it fails to predict the extent of thermal dissatisfaction of people in buildings.
The \ac{pmv} was free from bias only when it was used to predict thermal neutrality.
While, \mycite{Cheung2019} determined the accuracy of the \ac{pmv} model, by comparing its results with the \ac{tsv} from the ASHRAE Global Thermal Comfort Database II.
Finally, while \mycite{Yao2022} compare the \ac{pmv} and \gls{pmv-ce} models, their aim was primarily to compare these two formulations with other adaptive \ac{pmv} formulations, hence, they do not provide a detailed analysis on the prediction accuracy of the two models.
In particular, they focus significantly how the models perform in different climates and when applied to people from different world regions, but for example they fail to report how the \ac{pmv} accuracy varies as a function of \ac{tsv}.
Moreover, their analysis only reported the mean bias of the different models.
This as depicted by \mycite{Humphreys2000} does not provide sufficient insights and information in depicting which model is more accurate in predicting \ac{tsv} particularly because the ASHRAE Global Thermal Comfort Database II v2.1 is an unbalanced dataset.

\subsection{Aim and Objectives}\label{subsec:aim-and-objectives}
This study aims to specifically fill this knowledge gap by comparing the results of the \ac{pmv} and \gls{pmv-ce} models which are currently used in the \gls{7730} and \gls{55} Standards.
We compared the results against the self-reported \ac{tsv} contained in the ASHRAE Global Thermal Comfort Database II v2.1, referred to as the \gls{db2} thereafter.
Our results have significant and profound implications for both researchers, practitioners, and policymakers since it provides quantitative evidence on which formulation of the \ac{pmv} model is more accurate.
This in turn will allow them to take an informed decision on which model it should be used in National and International Standards to predict thermal sensation.
Policymakers would also benefit from our results since they can use them to review the current versions of the \gls{7730} and \gls{55} Standards.

>>>>

\section{Introduction}\label{sec:introduction}
In 1970, Fanger~\cite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} Standard~\cite{iso7730} in its original form.
% todo also mention the other standards that use the PMV The PMV index has been widely used in national and international standards and guidance ASHRAE 55 [4], EN 15251 (new version EN 16798 [5]) [6], ISO 7730 [1], CIBSE Guide A [7] and GB/T 50785 [110] as shown in  Table 5. This sentence is from Yao 2022.
The \ac{pmv} model has been widely used worldwide by researchers and practitioners.
It is the most extensively used thermal comfort index for estimating thermal sensation.
A search within the article title, abstract, and keyword on Scopus with the search term ``predicted mean vote'' returns approximately 1600 documents.
Out of those 1082 are scientific peer-reviewed articles published in the research area of engineering, environmental science, social sciences, and energy.
This highlights this model's extensive adoption and use among the scientific community.

To this date, the \ac{pmv} still remains to be the most utilized thermal comfort model even though several studies have highlighted that the \ac{pmv} has low accuracy in predicting thermal sensation votes~\cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988}.

To overcome the accuracy concerns, in particular when high air speed is used to cool people, the \gls{55} Standard uses a modified version of the original model developed by Fanger, the \ac{pmv-ce}~\cite{ashrae552023}. 
The best source to see how \ac{pmv-ce} works is the standard itself, but the standard does not explain why some choices were implemented. 
A partial justification of the model is described in \mycite{arens_moving_2009} and \mycite{yang_cooling_2015}. 
The process used to calculate the \ac{pmv-ce} in the \gls{55} is summarized in the flowchart depicted in Figure~\ref{fig:flowchart_pmv_ce}.
In summary, when the \ac{v} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
The value of \ac{ce} is calculated using the \ac{set} equation, and it is then subtracted from both the \ac{tdb} and \ac{tr}.
The resulting values become the new inputs in the \ac{pmv} model.
Since the \ac{ce} accounts for the convective heat losses from the person to its environment, thus, the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
The other three input parameters (\ac{clo}, \ac{met}, and \ac{rh}) remain unchanged.
Consequently, for a given thermal environment, results of the two \ac{pmv} formulations differ when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
The rationale for this approach is that the original formulation of the \ac{pmv} model cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v} is higher than \qty{0.1}{\m\per\s}.
However, this method has limitations.
First, it should be noted that over the years, the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
Second, calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the two-node model.
Third, we could not find any peer-reviewed scientific publication that quantified the accuracy improvements of this model over the original \ac{pmv}.

@@

        \draw [arrow] (start) -- (dec1);
        \draw [arrow] (dec1) -- node[above,pos=0.3] {Yes} (pro1a);
        \draw [arrow] (dec1) -- node[right,pos=0.3] {No} (dec2);
        \draw [arrow] (pro1a) -- ($(pro1a.south)+(0,-0.5)$) -| (dec2);
        \draw [arrow] (dec2) -- node[above,pos=0.3] {Yes} (pro2a);
        \draw [arrow] (dec2) -- node[right,pos=0.3] {No} (dec3);
        \draw [arrow] (pro2a) -- ($(pro2a.south)+(0,-0.5)$) -| (dec3);
        \draw [arrow] (dec3) -- ($(dec3.west)$) -| node[above,pos=0.3] {No} (end1);
        \draw [arrow] (dec3) -- ($(dec3.east)$) -| node[above,pos=0.3] {Yes} (pro3);
        \draw [arrow] (pro3) -- (end2);
    \end{tikzpicture}
    \caption{Flowchart depicting the steps for the calculation of the PMV in accordance with the \acs{55} Standard.}
    \label{fig:flowchart_pmv_ce}
\end{figure}

Choosing between the \ac{pmv} and \ac{pmv-ce} is consequently now a source of confusion for many practitioners, educators, and researchers worldwide since both models are widely adopted and incorporated in many building codes and standards worldwide.

Several other formulations of the \ac{pmv} model have been proposed to complicate this issue further.
Among the most notable formulations there are the: \ac{pmvs}~\cite{GaggeSET}, \ac{pmvg}~\cite{GaggeSET}, \ac{epmv}~\cite{Toftum2002}, and \ac{athb}~\cite{Schweiker2022}.
\mycite{Yao2022} provides a comprehensive review of different thermal comfort models and compares and describes some of the above-mentioned models.

%In 2002, Fanger, P. O. and Toftum, J. developed the \ac{epmv} an extension of the \ac{pmv}.
%The \ac{epmv} includes an expectancy factor introduced for use in non-air-conditioned buildings in warm climates.
%The authors argued that people adapt to the warm environment by decreasing their \ac{met} rate.
%They, consequently, propose to reduce the metabolic rate used to calculate the \ac{pmv} by \qty{6.7}{\percent} for every scale unit of \ac{pmv} above neutral.
%The resulting \ac{pmv} value is multiplied by the \ac{ef}.
%The \ac{ef} for non-air-conditioned buildings can be determined by the length of warm weather throughout the year and whether such buildings may be compared to many others in the region that are air-conditioned.
%If the temperature is hot all year or most of the year and there are no or few other air-conditioned buildings, \ac{ef} may be \num{0.5}, but if there are many other air-conditioned buildings, \ac{ef} may be \num{0.7}.
%For non-air-conditioned buildings in locations where the temperature is warm solely during the summer and no or few buildings have air-conditioning \ac{ef} is \numrange{0.7}{0.8}, whereas it is \numrange{0.8}{.9} in areas where there are numerous air-conditioned buildings.
%In regions with only brief periods of warm weather during the summer, the expectancy factor may be \numrange{0.9}{1}.
%This arbitrary and non-scientific definition of the expectancy factor make the \ac{epmv} difficult to calculate and non-generalizable.
%Consequently, the \ac{epmv} has not gained any widespread adoption.
%
%Yao, R. et al. (2009) developed the Adaptive Predicted Mean Vote Model~\gls{apmv} to account for occupants adaptation.
%They used a ``Black Box'' model to estimate an adaptive coefficient which is in turn use to calculate the \gls{apmv} using the following equation: $\mathrm{aPMV} = \mathrm{PMV}/(1+\lambda \mathrm{PMV})$
%The main issue with this approach is that the adaptive coefficient cannot be known a priori, but it is estimated based on the \ac{tsv} collected by a cohort of participants.
%This approach cannot be, therefore, generalizable and similarly to the \ac{epmv} model the \gls{apmv} is rarely used in thermal comfort research.
%

\subsection{Comparison of ISO and ASHRAE PMV formulations}\label{subsec:comparision-of-pmv-formulations}
To our knowledge, no previous study compared in detail the accuracy of the \ac{pmv} and \ac{pmv-ce} models, which are used in the \gls{7730} and \gls{55} Standards.

Some notable works that tried to determine the accuracy of the \ac{pmv} model include the manuscript from \mycite{doherty_evaluation_1988} who evaluated the ability of the \ac{pmv} and \ac{set} models to predict several physiological variables (i.e., skin temperature, core temperature, and skin wettedness) under a wide range of still air environments and metabolic rates.
They concluded that the \ac{pmv} model is accurate for simulations of resting subjects, but its accuracy decreases as a function of metabolic rate.
Humphreys and Nicol estimated the effects of measurement and formulation error on predicting thermal sensation using the \ac{pmv}~\cite{Humphreys2000}.
They used the ASHRAE Global Thermal Comfort database I and determined that the measurement error and the error introduced by the \ac{pmv} model formulation had a similar and non-negligible contribution.
They also determined the validity of the \ac{pmv} for predicting comfort votes collected in field studies~\cite{Humphreys2002}.
They concluded that the \ac{pmv} range of applicability should be significantly reduced and it fails to predict the extent of thermal dissatisfaction of people in buildings.
The \ac{pmv} was free from bias only when it was used to predict thermal neutrality.
\mycite{Cheung2019} determined the accuracy of the \ac{pmv} model, by comparing its results with the \ac{tsv} from the ASHRAE Global Thermal Comfort Database II.
They found that the thermal sensation predicted by the PMV model, on average, is one full thermal sensation scale unit away from the subject’s responses, confirming the results of~\cite{Humphreys2002}.
The accuracy of PMV was only \qty{34}{\percent}, the model has a slightly higher prediction accuracy for sensation close to neutrality, but the accuracy declined towards either end of the thermal sensation scale, and it overestimated both hot and cold sensations.
They also found that the Predicted Percentage of Dissatisfied (PPD) failed to predict the percentage of unacceptable votes if the thermal sensation was predicted using the PMV model and suggested its removal from \gls{55} Standards.
These results were confirmed based on the Chinese Thermal Comfort database~\cite{du_evaluation_2022}.
\mycite{Yao2022} compared the \ac{pmv} and \ac{pmv-ce} models, however, their aim was primarily to compare these two formulations with other adaptive \ac{pmv} formulations, hence, they do not provide a detailed analysis on the prediction accuracy of the two models.
In particular, they focus significantly on how the models perform in different climates and when applied to people from different world regions, but for example, they do not report how the \ac{pmv} accuracy varies as a function of environmental and personal parameters, and \ac{tsv}.
Their analysis only reports the mean bias of the different models.
This, as depicted by \mycite{Humphreys2000}, does not provide sufficient insights and information in determining which model is more accurate in predicting \ac{tsv}.
This is because the \ac{pmv} model has a higher prediction accuracy of sensation close to neutrality. 
Being the ASHRAE Global Thermal Comfort Database II v2.1 an unbalanced dataset and most of the data points are neutral a more detailed analysis as the one we are presenting in this manuscript is needed to compare the accuracy of both formulations.

\subsection{Aim and Objectives}\label{subsec:aim-and-objectives}
We aim to compare the accuracy of the \ac{pmv} and \ac{pmv-ce} models used in the \gls{7730} and \gls{55} Standards.
We used the self-reported \ac{tsv} recorded in the \ac{db2}.
We aim to determine which \ac{pmv} model is most accurate, this will allow policymakers to make an informed decision on which model should be used in standards.
